### 2022-02-18 paper/output/document/thesis_2_methods.tex
started chapter 4

Most theories of the good require that a moral reasoner understand complex features about
individuals' preferences, desires, or sensations in order to evaluate a moral action, making automated
consequentialist ethics difficult. Evaluating a state of affairs requires many controversial
judgements about whether a state of affairs actually satisifes the relevant criteria for goodness. 
Perfect knowledge of tens of thousands of 
people's pleasure or preferences or welfare or rights is impossible. Either a human being 
assigns values to states of affairs, which doesn't scale, or the machine does, 
which requires massive common-sense and increases room for doubting the system's judgements. This may be 
a tractable problem, but it is much more difficult than the equivalent Kantian task of formulating
and evaluating an action.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\noindent \textbf{Aggregation}%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
Once an automated consequentialist agent assigns a goodness measurement to each person in a state of affairs, it 
must also calculate an overall goodness measurement for the state of affairs. One approach to assigning
this value is to aggregate each person's individual goodness score into one complete score for a state. 
The more complex the theory of the good, the more difficult this aggregation becomes. For example, 
pluralistic theories struggle to explain how different kinds of value can be compared \citep{consequentialismsep}. 
How do we compare one unit of beauty to one unit of pleasure? Resolving this debate requires that an automated reasoner 
choose one specific aggregation algorithm, but those who disagree with this choice will not trust 
the reasoner's moral judgements. Moreover, for complex theories of the good, this aggregation algorithm
may be complex and may require a lot of data.

@@

ethical agent will need to commit to a particular theory of the virtues, a controversial choice. 

@@

those that an ideal virtuous agent would undertake. These 

@@

Just as prior work in consequentialism takes implicit or explicit stances on debates in consequentialist
literature, so does work in machine learning-based virtue ethics. For example, the training 
dataset with acts labelled as virtuous or not virtuous will contain an implicit view on what the virtues
are and how certain acts impact an agent's moral character. Because there is no canonical list of virtues
that virtue ethicists accept, this implicit view will likely be controversial.

Machine learning approaches like the Delphi system \citep{delphi} mentioned in Chapter \ref{intro} also may suffer explanability 
problems that my logic-programming, theorem-prover
approach does not face. Many machine learning algorithms cannot sufficiently explain their 
decisions to a human being, and often find patterns in datasets that don't 
cohere with the causes that a human being would identify \citep{puiutta}. While there is significant activity 
and progress in explainable machine learning, interactive theorem provers are designed to be explainable 
at the outset. Isabelle can show the axioms and lemmas it used in constructing a proof, 
allowing a human being to reconstruct the proof independently if they wish. This is not an 
intractable problem for machine learning approaches to computational ethics, but is one reason to 
prefer logical approaches.\footnote{This argument about explanability is in the context of virtue ethics and 
machine learning. It also applies to a broader class of work in automated ethics 
that uses ``bottom-up" approaches, in which a system learns moral judgements from prior judgements. 
I will extend this argument to general bottom-up approaches in Section Related Work.}%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Kantian Ethics \label{kantianethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this paper I focus on Kantian ethics. Kant's theory is centered 
on practical reason, which is the kind of reason that we 
use to decide what to do and the source of our agency. In \emph{The Groundwork of the Metaphysics of Morals}, Kant explains that 
rational beings are unique because we act ``in accordance with 
the representations of laws'' \citep[4:412]{groundwork}. In contrast, a ball thrown into the air acts 
according to the laws of physics. It cannot ask itself, ``Should I fall back to the ground?'' 
It simply falls. A rational being, on the other hand, can ask, ``Should I act on this reason?"''
As Korsgaard describes it, when choosing which desire to act on, ``it is as if there is something over 
and above all of your desires, something which is you, and which chooses which desire to act on'' \citep[100]{sources}. 
Rational beings are set apart by this reflective capacity. We are purposive and 
our actions are guided by practical reason. We have reasons for acting, even when these reasons are
opaque to us. This reflective choosing, or operation of practical reason, is what Kant calls the will.

@@

Armed with this understanding of practical reason, Kant presents the categorical 
imperative. He presents three ``formulations" or versions of the categorical imperative. In this project, 
I focus on the first formulation, the Formula of Universal Law, and I explain this choice in Section \ref{whyful}.

@@

Velleman presents a concise argument for the FUL. He argues that reason is universally shared among reasoners. For 
example, all reasoners have equal access to the arithmetic logic that shows that ``2+2=4'' \cite[29]{velleman}. The 
reasoning that makes this statement true is not specific to any person, but is universal across people. 
Therefore, if I have sufficient reason to will a maxim, so does every other rational agent. There is 
nothing special about the operation of my practical reason. 
In adopting a maxim, I implicitly state that all reasoners
across time also have reason to adopt that maxim. Therefore, because I act on reasons, I must obey the 
FUL. Notice that this fulfills the above criterion for a categorical imperative: the FUL is derived from 
a property of practical reason itself and thus derives authority from the will's authority over itself.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\noindent \textbf{Ease of Automation}%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
Kantian ethics is an especially candidate for formalization because the categorical imperative, particularly the FUL, 
is a property of reason related to the form or structure of a maxim. It does not require any situational 
knowledge beyond the circumstances included
in the maxim itself and thus requires fewer contingent facts than other ethical theories.
While other ethical theories often rely on many facts about 
the world or the actor, a computer evaluating 
a maxim doesn't require any knowledge about the world beyond what is contained in a maxim. Automating 
Kantian ethics merely requires making the notion of a maxim precise and representing it to the computer. 
This distinguishes Kantian ethics from consequentialism and virtue ethics, which
require far more knowledge about the world or the agent to reach a moral decision.

A maxim itself is an object with a thin representation for a computer, as compared to more complex 
objects like states of affairs or moral character. Later in my project, I argue that a maxim can be 
represented simply as a tuple of circumstances, act, and goal.\footnote{For more, see Section \ref{whatisamaxim}.} 
This representation is simple and efficient, especially when compared to the representations of a causal 
chain or a state of affairs or moral character. This property not only reduces the computational complexity
(in terms of time and space) of representing a maxim, but it also makes the system easier for human reasoners
to interact with. A person crafting an input to a Kantian automated agent needs to reason about relatively
simple units of evaluation, as opposed to the more complex features that consequentialism and virtue
ethics require.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\noindent \textbf{Difficulties in Automation}%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
One debate in Kantian 
ethics is the role of ``common-sense'' reasoning. Kantian ethics requires common-sense reasoning to 
determine which circumstances are ``morally relevant'' in the formulation of a maxim. Many misunderstandings
in Kantian ethics are due to badly formulated maxims, so this question is important for an ethical 
reasoner to answer.\footnote{For example, critics of Kantian ethics worry that the maxim, ``When I am a
man, I will marry a man because I want to spend my life with him'' fails the universalizability
test because if all men only married men, sexual reproduction would stop. This argument implies 
that Kantian ethics is homophobic. Kantians often respond by arguing that the correct formulation of 
this maxim is, ``When I love a man, I will marry him because I want to spend my life with him,'' which
is universalizable because if everyone marries who they love, some men will marry women and others will
marry men.} My system does not need to answer this question because I assume a properly crafted
maxim as input and apply the categorical imperative to this input. Using my system in a fully automated
moral agent will require answering this question, a challenging computational and philosophical task.

Common-sense reasoning is also relevant when applying the universalizability test itself. Consider the example
maxim ``When broke, I will falsely promise to repay a loan
to get some quick cash.'' This maxim fails the universalizability test because in a world where everyone
falsely promises to repay loans, no one will believe promises anymore, so the maxim will no longer serve
its intended purpose (getting some quick cash). Making this judgement requires understanding enough about
the system of promising to realize that it breaks down if everyone abuses it in this manner. This is a
kind of common sense reasoning that an automated Kantian agent would need. This need is not unique to
Kantian ethics; consequentialists agents need common sense to determine the consequences of 
an action and virtue ethical agents need common sense to determine which virtues an action
reflects. Making any ethical judgement requires robust conceptions of the action
at hand, falsely promising in this case. The advantage of Kantian ethics is that this is all the common 
sense that it requires, whereas a consequentialist or virtue ethical agent will require much more\footnote{
In Sections Lying and Murderer, I also use my system to demonstrate that Kantian ethics requires relatively thin conceptions 
of concepts like falsely promising.}. All
moral theories evaluating falsely promising will a robust definition of promising, 
but consequentialism and virtue ethics will also require additional information
that Kantian ethics will not. Thus, although the need for common sense poses a challenge to automated
Kantian ethics, this challenge is more acute for consequentialism or virtue ethics.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{The Formula of Universal Law \label{whyful}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Earlier I mentioned that Kant presents three formulations, or versions, 
of what he calls the ``supreme law of morality," but that I focus on the first of these three. In this section, 
I argue that the Formula of Universal Law, specifically, is the easiest part of Kantian ethics to automate
and the most generalizable.

@@

I choose to focus on the FUL\footnote{The FUL is often seen as emblematic of Kantian constructivism \cite[173]{ebelsduggan}. 
My project is not committed to Kantian constructivism. I believe that computational ethics is likely a 
valuable tool for any ethicist, and I make the case for Kantian ethics specifically.},
 because it is the most formal and thus the easiest to formalize and implement. 
Onora O'Neill explains that the formalism of the FUL allows 
for greater precision in philosophical arguments analyzing its implications and power \cite[33]{actingonprinciple}. This precision 
is particularly useful in a computational context because any formalism necessarily makes its content 
precise. The FUL's precision reduces ambiguity, allowing me to remain faithful to philosophical
literature on Kant. Precision reduces the need to make choices to resolve debates 
and ambiguities. Minimizing these choices minimizes 
arbitrariness in my formalization and puts it on solid philosophical footing.

@@

%:%184=115%:%
%:%188=117%:%

@@

%:%208=136%:%
%:%212=139%:%

@@

%:%236=157%:%
%:%248=159%:%

@@

%:%258=169%:%
%:%259=170%:%
%:%260=171%:%
%:%261=172%:%
%:%262=173%:%
%:%263=174%:%
%:%264=175%:%
%:%265=176%:%
%:%266=177%:%
%:%270=179%:%
%:%274=181%:%
%:%275=182%:%
%:%276=183%:%
%:%277=184%:%
%:%278=185%:%
%:%279=186%:%
%:%280=187%:%
%:%281=188%:%
%:%282=189%:%
%:%283=190%:%
%:%284=191%:%
%:%285=192%:%
%:%286=193%:%
%:%287=194%:%
%:%288=195%:%
%:%289=196%:%
%:%290=197%:%
%:%291=198%:%
%:%292=199%:%
%:%293=200%:%
%:%294=201%:%
%:%295=202%:%
%:%296=203%:%
%:%297=204%:%
%:%298=205%:%
%:%299=206%:%
%:%300=207%:%
%:%309=210%:%
%:%321=212%:%
%:%322=213%:%
%:%323=214%:%
%:%324=215%:%
%:%325=216%:%
%:%326=217%:%
%:%327=218%:%
%:%328=219%:%
%:%329=220%:%
%:%330=221%:%
%:%331=222%:%
%:%332=223%:%
%:%333=224%:%
%:%334=225%:%
%:%335=226%:%
%:%336=227%:%
%:%337=228%:%
%:%338=229%:%
%:%339=230%:%
%:%340=231%:%
%:%341=232%:%
%:%342=233%:%
%:%343=234%:%
%:%344=235%:%
%:%345=236%:%
%:%346=237%:%
%:%347=238%:%
%:%348=239%:%
%:%349=240%:%
%:%350=241%:%
%:%351=242%:%
%:%352=243%:%
%:%353=244%:%
%:%354=245%:%
%:%355=246%:%
%:%356=247%:%
%:%357=248%:%
%:%358=249%:%
%:%359=250%:%
%:%360=251%:%
%:%361=252%:%
%:%362=253%:%
%:%363=254%:%
%:%364=255%:%
%:%365=256%:%
%:%366=257%:%
%:%367=258%:%
%:%368=259%:%
%:%369=260%:%
%:%370=261%:%
%:%371=262%:%
%:%372=263%:%
%:%373=264%:%
%:%374=265%:%
%:%375=266%:%
%:%376=267%:%
%:%377=268%:%
%:%381=271%:%
%:%385=273%:%
%:%386=274%:%
%:%387=275%:%
%:%388=276%:%
%:%389=277%:%
%:%390=278%:%
%:%391=279%:%
%:%392=280%:%
%:%393=281%:%
%:%394=282%:%
%:%395=283%:%
%:%396=284%:%
%:%397=285%:%
%:%398=286%:%
%:%399=287%:%
%:%400=288%:%
%:%401=289%:%
%:%402=290%:%
%:%403=291%:%
%:%404=292%:%
%:%408=295%:%
%:%412=297%:%
%:%413=298%:%
%:%414=299%:%
%:%415=300%:%
%:%416=301%:%
%:%417=302%:%
%:%418=303%:%
%:%419=304%:%
%:%420=305%:%
%:%421=306%:%
%:%422=307%:%
%:%423=308%:%
%:%424=309%:%
%:%425=310%:%
%:%426=311%:%
%:%427=312%:%
%:%428=313%:%
%:%429=314%:%
%:%430=315%:%
%:%431=316%:%
%:%432=317%:%
%:%433=318%:%
%:%434=319%:%
%:%435=320%:%
%:%436=321%:%
%:%437=322%:%
%:%438=323%:%
%:%439=324%:%
%:%440=325%:%
%:%441=326%:%
%:%442=327%:%
%:%443=328%:%
%:%452=330%:%
%:%464=333%:%
%:%465=334%:%
%:%466=335%:%
%:%467=336%:%
%:%468=337%:%
%:%469=338%:%
%:%470=339%:%
%:%471=340%:%
%:%472=341%:%
%:%473=342%:%
%:%474=343%:%
%:%475=344%:%
%:%476=345%:%
%:%477=346%:%
%:%478=347%:%
%:%479=348%:%
%:%480=349%:%
%:%481=350%:%
%:%482=351%:%
%:%483=352%:%
%:%484=353%:%
%:%485=354%:%
%:%486=355%:%
%:%487=356%:%
%:%488=357%:%
%:%489=358%:%
%:%490=359%:%
%:%491=360%:%
%:%492=361%:%
%:%493=362%:%
%:%494=363%:%
%:%495=364%:%
%:%496=365%:%
%:%497=366%:%
%:%498=367%:%
%:%499=368%:%
%:%500=369%:%
%:%501=370%:%
%:%502=371%:%
%:%503=372%:%
%:%504=373%:%
%:%505=374%:%
%:%506=375%:%
%:%507=376%:%
%:%508=377%:%
%:%509=378%:%
%:%510=379%:%
%:%511=380%:%
%:%512=381%:%
%:%513=382%:%
%:%514=383%:%
%:%523=385%:%
%:%535=387%:%
%:%536=388%:%

@@

%:%541=392%:%
%:%542=393%:%
%:%543=394%:%
%:%544=395%:%
%:%545=396%:%
%:%546=397%:%
%:%547=398%:%
%:%548=399%:%
%:%549=400%:%
%:%550=401%:%
%:%551=402%:%
%:%552=403%:%
%:%553=404%:%
%:%554=405%:%
%:%555=406%:%
%:%556=407%:%
%:%557=408%:%
%:%558=409%:%
%:%559=410%:%
%:%560=411%:%
%:%561=412%:%
%:%562=413%:%
%:%563=414%:%
%:%564=415%:%
%:%565=416%:%
%:%566=417%:%
%:%567=418%:%
%:%571=420%:%
%:%572=421%:%
%:%573=422%:%
%:%574=423%:%
%:%575=424%:%
%:%576=425%:%
%:%577=426%:%
%:%578=427%:%
%:%579=428%:%
%:%580=429%:%
%:%581=430%:%
%:%582=431%:%
%:%583=432%:%
%:%584=433%:%
%:%585=434%:%
%:%586=435%:%
%:%587=436%:%
%:%588=437%:%
%:%589=438%:%
%:%590=439%:%
%:%591=440%:%
%:%592=441%:%
%:%593=442%:%
%:%594=443%:%
%:%595=444%:%
%:%596=445%:%
%:%597=446%:%
%:%606=449%:%
%:%618=451%:%
%:%619=452%:%
%:%620=453%:%
%:%621=454%:%
%:%622=455%:%
%:%623=456%:%
%:%632=459%:%
%:%644=461%:%
%:%645=462%:%
%:%646=463%:%
%:%648=465%:%

@@

%:%650=465%:%
%:%651=465%:%
%:%653=466%:%
%:%654=467%:%
%:%655=467%:%
%:%656=468%:%

@@

%:%658=469%:%
%:%659=469%:%
%:%660=469%:%
%:%662=470%:%
%:%663=471%:%
%:%664=471%:%
%:%667=485%:%
%:%668=486%:%
%:%669=487%:%
%:%671=489%:%

@@

%:%673=489%:%
%:%674=489%:%
%:%675=490%:%
%:%676=490%:%
%:%677=490%:%
%:%680=494%:%
%:%681=495%:%

>>>>

Most theories of the good require that a moral reasoner understand complex features about
individuals' preferences, desires, or sensations in order to evaluate a moral action, making automated
consequentialist ethics difficult. Evaluating a state of affairs requires many controversial
judgements about whether a state of affairs actually satisifes the relevant criteria for goodness. 
Perfect knowledge of tens of thousands of people's pleasure or preferences or welfare or rights is 
difficult, if not impossible.\footnote{Even if it were possible, collecting this kind of data poses 
privacy and surveillance risks.} Either a human being 
assigns values to states of affairs, which doesn't scale, or the machine does, 
which requires massive common-sense and increases room for doubting the system's judgements. This may be 
a tractable problem, but it is much more difficult than the equivalent Kantian task of formulating
and evaluating an action.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\noindent \textbf{Aggregation}%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
Once an automated consequentialist agent assigns a goodness measurement to each person in a state of affairs, it 
must also calculate an overall goodness measurement for the state of affairs. One approach to assigning
this value is to aggregate each person's individual goodness score into one complete score for a state. 
The more complex the theory of the good, the more difficult this aggregation becomes. For example, 
pluralistic theories struggle to explain how different kinds of value can be compared \citep{consequentialismsep}. 
How do we compare one unit of beauty to one unit of pleasure? Resolving this debate requires that an automated reasoner 
choose one specific aggregation algorithm, but those who disagree with this choice will not trust 
the reasoner's moral judgements. Moreover, for complex theories of the good, this aggregation algorithm
may be complex and may require a lot of data.

@@

ethical agent will need to commit to a list and definitions of the virtues, a controversial choice. 

@@

those that an ideally virtuous agent would undertake. These 

@@

Just as prior work in consequentialism takes implicit or explicit stances on debates in consequentialist
literature, so does work in machine learning-based virtue ethics. For example, the training 
dataset with acts labelled as virtuous or not virtuous will contain an implicit view on what the virtues
are and how certain acts impact an agent's moral character. Because there is no canonical list of all virtues
that virtue ethicists accept, this implicit view will likely be controversial. Even virtue ethicists agree
that certain traits, like courage, are virtues debate the exact definitions of these traits.

Machine learning approaches like the Delphi system \citep{delphi} mentioned in Chapter \ref{intro} also may suffer explanability 
problems that my logic-programming, theorem-prover
approach does not face. Many machine learning algorithms cannot sufficiently explain their 
decisions to a human being, and often find patterns in datasets that don't 
cohere with the causes that a human being would identify \citep{puiutta}. While there is significant activity 
and progress in explainable machine learning, interactive theorem provers are designed to be explainable 
at the outset. Isabelle can show the axioms and lemmas it used in constructing a proof, 
allowing a human being to reconstruct the proof independently if they wish. This is not an 
intractable problem for machine learning approaches to computational ethics, but is one reason to 
prefer logical approaches.\footnote{This argument about explanability is in the context of virtue ethics and 
machine learning. It also applies to a broader class of work in automated ethics 
that uses a ``bottom-up" approach, in which a system learns moral judgements from prior judgements. 
I will extend this argument to general bottom-up approaches in Section Related Work.}%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Kantian Ethics \label{kantianethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this paper I focus on Kantian ethics. Kant's theory is centered 
on practical reason, which is the kind of reason that we 
use to decide what to do and the source of our agency. In \emph{The Groundwork of the Metaphysics of Morals}, Kant explains that 
rational beings are unique because we act ``in accordance with 
the representations of laws'' \citep[4:412]{groundwork}. In contrast, a ball thrown into the air acts 
according to the laws of physics. It cannot ask itself, ``Should I fall back to the ground?'' 
It simply falls. A rational being, on the other hand, can ask, ``Should I act on this reason?"''
As Korsgaard describes it, when choosing which desire to act on, ``it is as if there is something over 
and above all of your desires, something which is you, and which chooses which desire to act on'' \citep[100]{sources}. 
Rational beings are set apart by this reflective capacity. We are purposive and 
our actions are guided by practical reason. We have reasons for acting, even when these reasons are
opaque to us. This reflective choosing, or operation of practical reason, is what Kant calls the will.

@@

Armed with this understanding of practical reason, Kant presents the categorical 
imperative. He presents three ``formulations" or versions of the categorical imperative. In this project, 
I focus on the first formulation, the Formula of Universal Law, and I justify this choice in Section \ref{whyful}.

@@

Velleman presents a concise argument for the FUL. He argues that reason is universally shared among reasoners. For 
example, all reasoners have equal access to the arithmetic logic that shows that ``2+2=4'' \cite[29]{velleman}. The 
reasoning that makes this statement true is not specific to any person, but is universal across people. 
Therefore, if I have sufficient reason to will a maxim, so does every other rational agent. There is 
nothing special about the operation of my practical reason. 
In adopting a maxim, I implicitly state that all reasoners
across time also have reason to adopt that maxim. Therefore, because I act on reasons, I must obey the 
FUL. Notice that this fulfills the above criterion for a categorical imperative: the FUL is derived from 
a property of practical reason itself and thus derives authority from the will's authority over itself.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\noindent \textbf{Ease of Automation}%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
Kantian ethics is an especially attractive candidate for formalization because the categorical imperative, particularly the FUL, 
is a property of reason related to the form or structure of a maxim. It does not require any situational 
knowledge beyond the circumstances included
in the maxim itself and thus requires fewer contingent facts than other ethical theories.
While other ethical theories often rely on many facts about 
the world or the actor, a computer evaluating 
a maxim doesn't require any knowledge about the world beyond what is contained in a maxim. Automating 
Kantian ethics merely requires making the notion of a maxim precise and representing it to the computer. 
This distinguishes Kantian ethics from consequentialism and virtue ethics, which
require far more knowledge about the world or the agent to reach a moral decision.

A maxim itself is an object with a thin representation for a computer, as compared to more complex 
objects like states of affairs or moral character. Later in my project, I argue that a maxim can be 
represented simply as a tuple of circumstances, act, and goal.\footnote{For more, see Section \ref{whatisamaxim}.} 
This representation is simple and efficient, especially when compared to the representations of a causal 
chain or a state of affairs or moral character. This property not only reduces the computational complexity
(in terms of time and space) of representing a maxim, but it also makes the system easier for human reasoners
to interact with. A person crafting an input to a Kantian automated agent needs to reason about relatively
simple units of evaluation, as opposed to the more complex features that consequentialism and virtue
ethics require.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\noindent \textbf{Difficulties in Automation}%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
One challenge for automating Kantian ethics is the need for ``common-sense'', or factual and 
situational background. Common-sense is needed when formulating a maxim and when determining if a maxim 
violates the Formula of Universal Law. Maxims include the circumstances in which they apply and
determining which circumstances are ``morally relevant'' to a maxim requires factual background. 
Many misunderstandings in Kantian ethics are due to badly formulated maxims, so this question is 
important for an ethical reasoner to answer.\footnote{For example, critics of Kantian ethics worry 
that the maxim, ``When I am a
man, I will marry a man because I want to spend my life with him'' fails the universalizability
test because if all men only married men, sexual reproduction would stop. This argument implies 
that Kantian ethics is homophobic. Kantians often respond by arguing that the correct formulation of 
this maxim is, ``When I love a man, I will marry him because I want to spend my life with him,'' which
is universalizable because if everyone marries who they love, some men will marry women and others will
marry men.} My system does not need to answer this question because I assume a properly crafted
maxim as input and apply the categorical imperative to this input. Using my system in a fully automated
moral agent will require answering this question, a challenging computational and philosophical task.\footnote{I
discuss this problem in greater detail in Section \ref{whatisamaxim} and Section Limitations.}

Common sense is also relevant when applying the universalizability test itself. Consider the example
maxim ``When broke, I will falsely promise to repay a loan
to get some quick cash.'' This maxim fails the universalizability test because in a world where everyone
falsely promises to repay loans, no one will believe promises anymore, so the maxim will no longer serve
its intended purpose (getting some quick cash). Making this judgement requires understanding enough about
the system of promising to realize that it breaks down if everyone abuses it in this manner. This is a
kind of common sense reasoning that an automated Kantian agent would need. This need is not unique to
Kantian ethics; consequentialists agents need common sense to determine the consequences of 
an action and virtue ethical agents need common sense to determine which virtues an action
reflects. For example, in the case of virtue ethics, in order to see that saving a baby from a lion 
requires courage, a reasoner must have enough background knowledge to know that lions are scary. Making any 
ethical judgement requires robust conceptions of the action at hand. Kantian ethics requires far less common sense 
than a consequentialist or virtue ethical agent.\footnote{
In Sections Lying and Murderer, I also use my system to demonstrate that Kantian ethics requires relatively thin conceptions 
of concepts like falsely promising.} All
moral theories evaluating falsely promising will a robust definition of promising, 
but consequentialism and virtue ethics will require additional information
that Kantian ethics will not. Thus, although the need for common sense poses a challenge to automated
Kantian ethics, this challenge is more acute for consequentialism or virtue ethics.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{The Formula of Universal Law \label{whyful}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Earlier I mentioned that Kant presents three formulations, or versions, 
of what he calls the ``supreme law of morality," but that I focus on the first of these three. In this section, 
I argue that the Formula of Universal Law, specifically, is the easiest part of Kantian ethics to automate
and the most generalizable.

@@

I choose to focus on the FUL\footnote{The FUL is often seen as emblematic of Kantian constructivism \cite[173]{ebelsduggan}. 
My project is not committed to Kantian constructivism.}, because it is the most formal and thus the 
easiest to formalize and implement. Onora O'Neill explains that the formalism of the FUL allows 
for greater precision in philosophical arguments analyzing its implications and power \cite[33]{actingonprinciple}. This precision 
is particularly useful in a computational context because any formalism necessarily makes its content 
precise. The FUL's precision reduces ambiguity, allowing me to remain faithful to philosophical
literature on Kant. Precision reduces the need to make choices to resolve debates 
and ambiguities. Minimizing these choices minimizes 
arbitrariness in my formalization and puts it on solid philosophical footing.

@@

%:%181=114%:%
%:%185=116%:%

@@

%:%205=134%:%
%:%209=137%:%

@@

%:%228=155%:%
%:%237=158%:%

@@

%:%255=166%:%
%:%259=169%:%
%:%260=170%:%
%:%261=171%:%
%:%262=172%:%
%:%263=173%:%
%:%264=174%:%
%:%265=175%:%
%:%266=176%:%
%:%267=177%:%
%:%271=179%:%
%:%275=181%:%
%:%276=182%:%
%:%277=183%:%
%:%278=184%:%
%:%279=185%:%
%:%280=186%:%
%:%281=187%:%
%:%282=188%:%
%:%283=189%:%
%:%284=190%:%
%:%285=191%:%
%:%286=192%:%
%:%287=193%:%
%:%288=194%:%
%:%289=195%:%
%:%290=196%:%
%:%291=197%:%
%:%292=198%:%
%:%293=199%:%
%:%294=200%:%
%:%295=201%:%
%:%296=202%:%
%:%297=203%:%
%:%298=204%:%
%:%299=205%:%
%:%300=206%:%
%:%301=207%:%
%:%302=208%:%
%:%311=211%:%
%:%323=213%:%
%:%324=214%:%
%:%325=215%:%
%:%326=216%:%
%:%327=217%:%
%:%328=218%:%
%:%329=219%:%
%:%330=220%:%
%:%331=221%:%
%:%332=222%:%
%:%333=223%:%
%:%334=224%:%
%:%335=225%:%
%:%336=226%:%
%:%337=227%:%
%:%338=228%:%
%:%339=229%:%
%:%340=230%:%
%:%341=231%:%
%:%342=232%:%
%:%343=233%:%
%:%344=234%:%
%:%345=235%:%
%:%346=236%:%
%:%347=237%:%
%:%348=238%:%
%:%349=239%:%
%:%350=240%:%
%:%351=241%:%
%:%352=242%:%
%:%353=243%:%
%:%354=244%:%
%:%355=245%:%
%:%356=246%:%
%:%357=247%:%
%:%358=248%:%
%:%359=249%:%
%:%360=250%:%
%:%361=251%:%
%:%362=252%:%
%:%363=253%:%
%:%364=254%:%
%:%365=255%:%
%:%366=256%:%
%:%367=257%:%
%:%368=258%:%
%:%369=259%:%
%:%370=260%:%
%:%371=261%:%
%:%372=262%:%
%:%373=263%:%
%:%374=264%:%
%:%375=265%:%
%:%376=266%:%
%:%377=267%:%
%:%378=268%:%
%:%379=269%:%
%:%383=272%:%
%:%387=274%:%
%:%388=275%:%
%:%389=276%:%
%:%390=277%:%
%:%391=278%:%
%:%392=279%:%
%:%393=280%:%
%:%394=281%:%
%:%395=282%:%
%:%396=283%:%
%:%397=284%:%
%:%398=285%:%
%:%399=286%:%
%:%400=287%:%
%:%401=288%:%
%:%402=289%:%
%:%403=290%:%
%:%404=291%:%
%:%405=292%:%
%:%406=293%:%
%:%410=296%:%
%:%414=298%:%
%:%415=299%:%
%:%416=300%:%
%:%417=301%:%
%:%418=302%:%
%:%419=303%:%
%:%420=304%:%
%:%421=305%:%
%:%422=306%:%
%:%423=307%:%
%:%424=308%:%
%:%425=309%:%
%:%426=310%:%
%:%427=311%:%
%:%428=312%:%
%:%429=313%:%
%:%430=314%:%
%:%431=315%:%
%:%432=316%:%
%:%433=317%:%
%:%434=318%:%
%:%435=319%:%
%:%436=320%:%
%:%437=321%:%
%:%438=322%:%
%:%439=323%:%
%:%440=324%:%
%:%441=325%:%
%:%442=326%:%
%:%443=327%:%
%:%444=328%:%
%:%445=329%:%
%:%446=330%:%
%:%447=331%:%
%:%448=332%:%
%:%449=333%:%
%:%458=335%:%
%:%470=338%:%
%:%471=339%:%
%:%472=340%:%
%:%473=341%:%
%:%474=342%:%
%:%475=343%:%
%:%476=344%:%
%:%477=345%:%
%:%478=346%:%
%:%479=347%:%
%:%480=348%:%
%:%481=349%:%
%:%482=350%:%
%:%483=351%:%
%:%484=352%:%
%:%485=353%:%
%:%486=354%:%
%:%487=355%:%
%:%488=356%:%
%:%489=357%:%
%:%490=358%:%
%:%491=359%:%
%:%492=360%:%
%:%493=361%:%
%:%494=362%:%
%:%495=363%:%
%:%496=364%:%
%:%497=365%:%
%:%498=366%:%
%:%499=367%:%
%:%500=368%:%
%:%501=369%:%
%:%502=370%:%
%:%503=371%:%
%:%504=372%:%
%:%505=373%:%
%:%506=374%:%
%:%507=375%:%
%:%508=376%:%
%:%509=377%:%
%:%510=378%:%
%:%511=379%:%
%:%512=380%:%
%:%513=381%:%
%:%514=382%:%
%:%515=383%:%
%:%516=384%:%
%:%517=385%:%
%:%518=386%:%
%:%527=388%:%
%:%539=390%:%

@@

%:%544=394%:%
%:%545=395%:%
%:%546=396%:%
%:%547=397%:%
%:%548=398%:%
%:%549=399%:%
%:%550=400%:%
%:%551=401%:%
%:%552=402%:%
%:%553=403%:%
%:%554=404%:%
%:%555=405%:%
%:%556=406%:%
%:%557=407%:%
%:%558=408%:%
%:%559=409%:%
%:%560=410%:%
%:%561=411%:%
%:%562=412%:%
%:%563=413%:%
%:%564=414%:%
%:%565=415%:%
%:%566=416%:%
%:%567=417%:%
%:%568=418%:%
%:%569=419%:%
%:%570=420%:%
%:%571=421%:%
%:%575=423%:%
%:%576=424%:%
%:%577=425%:%
%:%578=426%:%
%:%579=427%:%
%:%580=428%:%
%:%581=429%:%
%:%582=430%:%
%:%583=431%:%
%:%584=432%:%
%:%585=433%:%
%:%586=434%:%
%:%587=435%:%
%:%588=436%:%
%:%589=437%:%
%:%590=438%:%
%:%591=439%:%
%:%592=440%:%
%:%593=441%:%
%:%594=442%:%
%:%595=443%:%
%:%596=444%:%
%:%597=445%:%
%:%598=446%:%
%:%599=447%:%
%:%600=448%:%
%:%601=449%:%
%:%610=452%:%
%:%622=454%:%
%:%623=455%:%
%:%624=456%:%
%:%625=457%:%
%:%626=458%:%
%:%627=459%:%
%:%636=462%:%
%:%648=464%:%

@@

%:%650=466%:%
%:%652=468%:%
%:%653=468%:%
%:%654=468%:%
%:%655=468%:%

@@

%:%658=470%:%
%:%659=470%:%
%:%660=471%:%
%:%661=472%:%
%:%662=472%:%
%:%663=472%:%
%:%664=472%:%
%:%666=473%:%
%:%667=474%:%
%:%668=474%:%
%:%671=488%:%

@@

%:%673=490%:%
%:%675=492%:%
%:%676=492%:%
%:%677=492%:%
%:%678=492%:%
%:%679=493%:%
%:%680=493%:%
%:%681=493%:%
